---
title: "*Phytophthora* hosts in UK forests and commercial forestry"
author: ""
date: ""
output: html_document

---

###Selecting resilient hosts

This tool is a searchable database of known *Phytophthora* pathogens of tree species used in UK forests and forestry. You can search by tree host, by pathogen, or by country.   

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = getwd())
library(DT)
library(raster)
library(knitr)
```
```{r host_data}

load("data/hosts.rData")



```

**Search for a UK host tree species to find associated *Phytophthora* pathogens**  
  
  
```{r host1, echo=FALSE, out.width = '50%'}

hosts <- hosts[,c("UK host species", "host family", "Phytophthora species", "Phytophthora present in UK", "country", "year", "reference" )]
datatable(hosts,
          rownames = FALSE, 
          escape = FALSE)




```
###Feedback

Thank you for taking the time to explore this data. Please follow the link below to complete your feedback. Your responses will be used to understand the value of information about which hosts are known to be susceptible to *Phytophthora* species in supporting plant health awareness in planting and procurement decisions for environmental and landscaping purposes.

*LB: It may be possible for us to embed a Google Forms item within the page (see https://github.com/ramnathv/RGoogleForms). However, we may need something more flexible for our questionnaire, in which case a link openeing in a new tab might work better. This is just a mock-up questionnaire...*

## [Enter my feedback - *Phytophthora* associated with hosts in UK forest and forestry?](<https://docs.google.com/forms/d/182WIK4SZFv0hQu9YyJ0FOddkD0I5w-HYJWm50hRDcyU/edit?usp=sharing>)