---
title: Where are new pests and pathogens coming from? 
author: "Louise Barwell, Bethan Purse and Daniel Chapman"
date: ""
output: html_document
---

```{r setup, echo = FALSE}

knitr::opts_chunk$set(echo = FALSE)

```

```{r include=FALSE}
library(leaflet)
library(RColorBrewer)
library(readr)
library(countrycode)
library(dplyr)
library(geosphere)
library(rgeos)
library(reshape2)
library(htmlTable)
library(rgdal)
library(sp)
library(leafem)
library(raster)
library(htmlwidgets)



load("data/export_risk_composite.rData")



# get a nice world map - someone recommeded this one:
# download.file(file.path('http://www.naturalearthdata.com/http/',
#                         'www.naturalearthdata.com/download/50m/cultural',
#                         'ne_50m_admin_0_countries.zip'), 
#               f <- tempfile())

#unzip(f, exdir=tempdir())

world <- readOGR(dsn = 'data', layer = 'ne_50m_admin_0_countries', encoding='UTF-8')
export_risk_composite$GU_A3 <- export_risk_composite$exporter_iso3
export_risk_composite <- export_risk_composite[export_risk_composite$importer_iso3 %in% world$GU_A3,]
world <- merge(world, export_risk_composite, all.x= TRUE) # 

# where no Phytophthora are reported (0) convert richness to NA 
world@data$Phytophthora_richness[world@data$Phytophthora_richness == 0] <- NA

pal_richness <- colorNumeric("YlOrRd", domain = log(world@data$Phytophthora_richness))
pal_risk <- colorNumeric("YlOrRd", domain = world@data$composite_export_risk)
pal_climate <- colorNumeric("YlOrRd", domain = log(world@data$climate_raw))
pal_trade <- colorNumeric("YlOrRd", domain = log(1+world@data$trade_raw))
pal_expproactive <- colorNumeric("YlOrRd", domain = world@data$expproactive_raw)

# EPPO data
load("data/EPPO_map_data.rData")
world2 <- readOGR(dsn = 'data', layer = 'ne_50m_admin_0_countries', encoding='UTF-8')
EPPO_map_data <- EPPO_map_data[EPPO_map_data$GU_A3 %in% world2$GU_A3,]
world2 <- merge(world2, EPPO_map_data, all.x= TRUE) # 

# where no Phytophthora are reported (0) convert richness to NA 
world2@data$richness[world@data$richness == 0] <- NA

pal_pest_richness <- colorNumeric("YlOrRd", domain = log(world2@data$richness))


```

###Global trade networks and recent invasions of non-native pests, pathogens and plants

This map allows you to explore the global source distributions of 422 non-native plant pests of concern within the EPPO (European Plant Protection organisation) region. Among these insect pests (173), pathogens (166) and weeds (83), live plant imports from nearby countries provided the best explanation of recent invasions (Chapman, D, Purse, BV, Roy, HE, Bullock, JM. Global trade networks determine the distribution of invasive non‐native species. Global Ecol Biogeogr. 2017; 26: 907– 917. https://doi.org/10.1111/geb.12599).

```{r EPPO_source, warning=FALSE, message=FALSE}



leaflet(data = world2, width = "100%") %>% addTiles(options = providerTileOptions(minZoom=1, maxZoom=10)) %>% 
  addLegend(pal = pal_pest_richness, 
            values = ~log(richness),
            bins = 6,
            labFormat = labelFormat(transform = function(x) exp(x), digits = 0),
            position = "topright",
            title = "Pest richness") %>%
  addPolygons(
    fillColor = ~pal_pest_richness(log(world2@data$richness)),
    weight = 1,
    opacity = 1,
    color = "grey",
    dashArray = "3",
    fillOpacity = 0.7,
    popup = ~species_table,
    popupOptions = c(maxWidth = 800),
    highlight = highlightOptions(
      weight = 1.5,
      color = "#666",
      dashArray = "",
      fillOpacity = 0.7,
      bringToFront = TRUE)
  ) %>%
  addHomeButton(extent(world2), layer.name = "Reset") 

```

###Import risk factors: a *Phytophthora* case study

International trade in live plants is considered the main pathway for the introduction of plant pests and pathogens species to new geographic regions. Climate matching between the exporting and importing regions and the biosecurity capacity of the importing and exporting nations may also influence how pests and pathogens spread globally. Pest reporting activity also varies among countries and will affect how well this spread can be predicted. We have used these risk factors to help predict *Phytophthora* import risk to the UK from exporting countries. We have also used the thermal tolerance ranges of these *Phytophthora* species to help predict which *Phytophthora* pathogens may be more likely to arrive and establish given the level of imports of live plants from the known source countries of that pathogen. 
  


```{r map_flows, warning=FALSE, message=FALSE}
# make the interactive map

#"annual exports to UK (metric tons of live plants)"

  leaflet(data = world, width = "100%") %>% addTiles(options = providerTileOptions(minZoom=1, maxZoom=10)) %>% 
  addPolygons(
    fillColor = ~pal_risk(world@data$composite_export_risk),
    group = "export risk (probability of new arrival in UK)",
    weight = 1,
    opacity = 1,
    color = "grey",
    dashArray = "3",
    fillOpacity = 0.7,
    popup = ~species_table,
    popupOptions = c(maxWidth = 600),
    highlight = highlightOptions(
      weight = 1.5,
      color = "#666",
      dashArray = "",
      fillOpacity = 0.7,
      bringToFront = TRUE)
  ) %>%
  addPolygons(
    fillColor = ~pal_richness(log(world@data$Phytophthora_richness)),
    group = "Phytophthora species richness",
    weight = 1,
    opacity = 1,
    color = "grey",
    dashArray = "3",
    fillOpacity = 0.7,
    popup = ~species_table,
    popupOptions = c(maxWidth = 600),
    highlight = highlightOptions(
      weight = 1.5,
      color = "#666",
      dashArray = "",
      fillOpacity = 0.7,
      bringToFront = TRUE)
  ) %>%
  addPolygons(
    fillColor = ~pal_trade(log(1+world@data$trade_raw)),
    group = "annual exports to UK (metric tons of live plants)",
    weight = 1,
    opacity = 1,
    color = "grey",
    dashArray = "3",
    fillOpacity = 0.7,
    popup = ~species_table,
    popupOptions = c(maxWidth = 600),
    highlight = highlightOptions(
      weight = 1.5,
      color = "#666",
      dashArray = "",
      fillOpacity = 0.7,
      bringToFront = TRUE)
  ) %>% 
  addPolygons(
    fillColor = ~pal_climate(log(world@data$climate_raw)),
    group = "climate similarity (1/Mahalanobis distance)",
   weight = 1,
    opacity = 1,
    color = "grey",
    dashArray = "3",
    fillOpacity = 0.7,
    popup = ~species_table, 
   popupOptions = c(maxWidth = 600),
    highlight = highlightOptions(
      weight = 1.5,
      color = "#666",
      dashArray = "",
      fillOpacity = 0.7,
      bringToFront = TRUE)
   )  %>% 
  addPolygons(
    fillColor = ~pal_expproactive(world@data$expproactive_raw),
    group = "exporter biosecurity capacity",
    weight = 1,
    opacity = 1,
    color = "grey",
    dashArray = "3",
    fillOpacity = 0.7,
    popup = ~species_table,
    popupOptions = c(maxWidth = 600),
    highlight = highlightOptions(
      weight = 1.5,
      color = "#666",
      dashArray = "",
      fillOpacity = 0.7,
      bringToFront = TRUE)
  ) %>%
  addLegend(pal = pal_risk, 
            values = ~composite_export_risk,
            group = "export risk (probability of new arrival in UK)",
            position = "bottomleft",
            title = "export risk (probability of new arrival in UK)") %>% 
  addLegend(pal = pal_richness, 
            values = ~log(Phytophthora_richness),
            bins = 6,
            labFormat = labelFormat(transform = function(x) exp(x), digits = 0),
            group = "Phytophthora species richness",
            position = "bottomleft",
            title = "Phytophthora species richness") %>% 
  addLegend(pal = pal_trade, 
            values = ~log(1+trade_raw),
            labFormat = labelFormat(transform = function(x) exp(x) - 1, digits = 0),
            group = "annual exports to UK (metric tons of live plants)",
            position = "bottomleft",
            title = "annual exports to UK (metric tons of live plants)") %>% 
  addLegend(pal = pal_climate, 
            values = ~log(climate_raw), 
            labFormat = labelFormat(transform = function(x) exp(x), digits = 3),
            group = "climate similarity (1/Mahalanobis distance)",
            position = "bottomleft",
            title = "climate similarity (1/Mahalanobis distance)")  %>%
  addLegend(pal = pal_expproactive, 
            values = ~expproactive_raw, 
            group = "exporter biosecurity capacity",
            position = "bottomleft",
            title = "exporter biosecurity capacity") %>% 
  addLayersControl(
        baseGroups = c("export risk (probability of new arrival in UK)",
                       "Phytophthora species richness",
                       "annual exports to UK (metric tons of live plants)", 
                       "climate similarity (1/Mahalanobis distance)", 
                       "exporter biosecurity capacity"),
        options = layersControlOptions(collapsed = FALSE)
       ) %>% 
  # this section is a fudge to try to sync the legends with the layers shown when toggling between risk factors
  # amazingly, there is no inbuilt way to do this with baselayers (only with overlaid layers, where more than one layer can be shown at once)
htmlwidgets::onRender("
    function(el, x) {
      var updateLegend = function () {
          var selectedGroup = document.querySelectorAll('input:checked')[0].nextSibling.innerText.substr(1);

          document.querySelectorAll('.legend').forEach(a => a.hidden=true);
          document.querySelectorAll('.legend').forEach(l => {
            if (l.children[0].children[0].innerText == selectedGroup) l.hidden=false;
          });
      };
      updateLegend();
      this.on('baselayerchange', e => updateLegend());
    }") %>% 
  addHomeButton(extent(world), layer.name = "Reset") 

```




## [Enter my feedback - Where are new pests and pathogens coming from?](<https://forms.gle/YVwCV1E7v6LDwKBH7>)



